Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-02-13 ms.tex
Introduce x = cos iota much earlier

If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, \iota, \psi, \phi_c) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, inclination $\mathrm{\iota}$, polarization angle $\psi$, and coalescence phase $\phi_c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c) \right)^2
    \right]
\end{equation}
%
It is easily shown that
%
\begin{eqnarray*}
    {\rho_i}^2 &=
        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left[
            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
            \left(1 + 6 x^2 + x^4 \right) \\
            &+ \left(1 - x^2 \right) \left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi}
        \right],
\end{eqnarray*}
%
where $F_{+,i}$ and $F_{\times,i}$ are the `$+$' and `$\times$' antenna factors of detector $i$, themselves functions of sky location $\mathbf{n}$.

@@

Prior is uniform in $\mathbf{n}$, $t_\oplus$, $\cos \iota$, $\psi$, and $\phi_c$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume.

@@

The marginalization over $D_\mathrm{L}$, $\cos \iota$, and $psi$ must be done numerically. The outermost integration is over $\cos \iota$ and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $\cos \iota$ and $\psi$. This integration scheme is first order in $\cos \iota$ and second order in $\psi$.

>>>>

If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, x, \psi) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, x, \psi
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, cosine-inclination $x = \cos \iota$, and polarization angle $\psi$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, x, \psi) \right)^2
    \right]
\end{equation}
%
It is easily shown that
%
\begin{eqnarray*}
    {\rho_i}^2 &=
        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left[
            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
            \left(1 + 6 x^2 + x^4 \right) \\
            &+ \left(1 - x^2 \right) \left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi}
        \right],
\end{eqnarray*}
%
where $F_{+,i}$ and $F_{\times,i}$ are the `$+$' and `$\times$' antenna factors of detector $i$, themselves functions of sky location $\mathbf{n}$.

@@

Prior is uniform in $\mathbf{n}$, $t_\oplus$, x, and $\psi$. For distance, can choose either a prior that is uniform in $\log D_\mathrm{L}$ or uniform in volume.

@@

The marginalization over $D_\mathrm{L}$, $x$, and $psi$ must be done numerically. The outermost integration is over x and $\psi$, and is done with Newton-Cotes by evaluating $\mathcal{L}_\mathrm{SNR}$ on a regular lattice in $x$ and $\psi$. This integration scheme is first order in $x$ and second order in $\psi$.
